---
title: "Using blindspiker"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using blindspiker}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(blindspiker)
```

The R package, `blindspiker`, was initially developed for the  Internal Dosimetry program at the Savannah River Site, based on an R script, "SRS Blind Spikes-4.R", written by Tom Labone. 

The first step in using the package is to **load your data.** This is made easy **with the `get_data` ** function. This function reads in data directly from .csv files. 

Data must be supplied with column names indicated below:

Spike values must be in a comma separated variable file with the following column headers:

* sample_ID = unique identifier for the sample
* analyte = name of analyte in the sample
* spike_value = quantity of the analyte
* spike_unit = unit of measurement for the spike value
* provider_lab = for the spiked samples - Optional, not currently used for any function
* submission_date = date the sample was delivered to the laboratory in format yyyy-mm-dd (optional, not currently used, but user could consider it for later analysis of results in the combined data)
* sv_unc = uncertainty in the spike value - Optional, default = 0
* sv_k = coverage factor for the spike value uncertainty - Optional, default = 2


Laboratory values must be in a comma separated variable file with the following column headers:

* sample_ID = unique identifier for the sample - must match spike data
* analyte = name of analyte in the sample - must match spike data
* result_date = date the sample was reported in format yyyy-mm-dd
* k = coverage factor for the laboratory's uncertainty of the result - Optional, default = 2
* result = analysis result
* unc = analysis uncertainty in same units as result
* units = laboratory units - must match spike value units
* det_lvl = laboratory's detection level
      
Spike values provided to the laboratory may be summarized with the following functions:

**`table_spike` summarizes all individual spike values** and

**`spike_combos` provides any combinations** of spike values in individual samples that the user chooses to have tallied.

**`plot_run` provides run charts** with the option of plotting in analysis units or by a ratio of the laboratory results to the blind spike values. Both versions show uncertainty ranges of the results. Uncertainty bars are shown on the laboratory results and the ratios. When there is no overlap with the spike values, the laboratory result is either a false positive or a false negative. False positives are excluded from the ratio version when there is no spike value to avoid division by zero. 

**`table_false` provides confidence intervals on error rates** for all laboratory results. 

**`plot_tat` plots the laboratory turnaround time** for each spike sample.

**`plot_qq` provides quantile-quantile plots** of the results. Random errors are expected to fluctuate in a normally distributed pattern. When the QQ plot shows a deviation from normal, further investigation in the underlying process may be needed.
